Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-09-03 paper.tex
Fix figure caption

\subsection{Comparison of sampling patterns}
\label{sec:compare_patterns}
To compare different sampling and averaging patterns we performed computations
of $S_3$ on Sandstone 3D sample.  They were performed in three different
manners: 1) isosceles triangle with a right-triangle between the equal side --
this sampling resembles both our right-triangle pattern and pattern of Malmir et
al. \cite{malmir2018}, computations were made within XY planes within the image
and, thus, represent directional CF; 2) equilateral triangle with side $r$
directionally analyzed within XY planes; 3) equilateral triangle with side $r$
averaged over 20 random rotations -- reminiscence of weblike and random
patterns.

The results of the comparison are shown in \cref{fig:comparison_patterns}. For
isotropic samples curves for directional and ensemble averaged, as well as for
right-angle isosceles triangle, basically collapse to each other with minimal
differences.
\begin{figure*}[ht]
  \centering
  \subfigure[Isosceles with right angle]{
    \includegraphics[width=0.4\linewidth]{images/isosceles.png}}
  \hfill
  \subfigure[Equilateral]{
    \includegraphics[width=0.4\linewidth]{images/equilateral.png}}
  \vskip\baselineskip
  \subfigure[With random rotations]{
    \includegraphics[width=0.4\linewidth]{images/equilateral-avg.png}}
  \hfill
  \subfigure[Comparison plot]{
    \includegraphics[width=0.4\linewidth]{images/alltogether.png}}
  \caption[]{A comparison of computations of $S_3$ correlation function for the
    Sandstone sample using different patterns with and without random
    rotations. Left: evaluated with isosceles triangle pattern with the side $r$
    and the angle between edges $\phi$; sampling triangles are aligned with the
    plane XY.  Middle: evaluated with equilateral triangle pattern with the side
    $r$; sampling triangles are aligned with the plane XY.  Right: evaluated
    with equilateral triangle pattern with the side $r$. The result is averaged
    over 20 random rotations of the sampling triangle. Lower image: comparison
    of $S_3$ functions calculated with right triangle pattern with length of the
    both catheti $r$ and equilateral triangle pattern (both averaged over
    rotations and aligned with the plane XY).}
  \label{fig:comparison_patterns}
\end{figure*}

@@

Another interesting question is which sampling pattern is more powerful?
Our results shown in \cref{fig:comparison_patterns} clear illustrate that
for isotropic structure all sampling techniques provide quite similar results.
The differences, of course, will be significant for isotropic cases and
directional 3-point CFs calculations. Again, the clear winner here is not obvious.
Right-triangular pattern allows to systematically sample the structure with
predefined direction. On the other hand, equilateral triangle allows to
minimize the volume of computed functions and represent them as single curve,
similar to 2-point functions. Random pattern could be useful for rigorous bounds.
So, the better approach can depend on the problem at hand. Nonetheless,
the direct comparison of patterns can be performed only on the basis of
information content -- something we discuss in the next paragraph.

>>>>

\subsection{Comparison of sampling patterns}
\label{sec:compare-patterns}
To compare different sampling and averaging patterns we performed computations
of $S_3$ on Sandstone 3D sample.  They were performed in three different
manners: 1) isosceles triangle with a right-triangle between the equal side --
this sampling resembles both our right-triangle pattern and pattern of Malmir et
al. \cite{malmir2018}, computations were made within XY planes within the image
and, thus, represent directional CF; 2) equilateral triangle with side $r$
directionally analyzed within XY planes; 3) equilateral triangle with side $r$
averaged over 20 random rotations -- reminiscence of weblike and random
patterns.

The results of the comparison are shown in \cref{fig:comparison-patterns}. For
isotropic samples curves for directional and ensemble averaged, as well as for
right-angle isosceles triangle, basically collapse to each other with minimal
differences.
\begin{figure*}[ht]
  \centering
  \subfigure[Isosceles with right angle]{
    \includegraphics[width=0.4\linewidth]{images/isosceles.png}
    \label{fig:cp-isosceles}}
  \hfill
  \subfigure[Equilateral]{
    \includegraphics[width=0.4\linewidth]{images/equilateral.png}
    \label{fig:cp-equilateral}}
  \vskip\baselineskip
  \subfigure[With random rotations]{
    \includegraphics[width=0.4\linewidth]{images/equilateral-avg.png}
    \label{fig:cp-equilateral-avg}}
  \hfill
  \subfigure[Comparison plot]{
    \includegraphics[width=0.4\linewidth]{images/alltogether.png}
    \label{fig:cp-comp}}
  \caption[]{A comparison of computations of $S_3$ correlation function for the
    Sandstone sample using different patterns with and without random
    rotations. Fig.~\ref{fig:cp-isosceles}: evaluated with isosceles triangle
    pattern with the side $r$ and the angle between edges $\phi$; sampling
    triangles are aligned with the plane XY. Fig.~\ref{fig:cp-equilateral}:
    evaluated with equilateral triangle pattern with the side $r$; sampling
    triangles are aligned with the plane XY. Fig.~\ref{fig:cp-equilateral-avg}:
    evaluated with equilateral triangle pattern with the side $r$. The result is
    averaged over 20 random rotations of the sampling
    triangle. Fig.~\ref{fig:cp-comp}: comparison of $S_3$ functions calculated
    with right triangle pattern with length of the both catheti $r$ and
    equilateral triangle pattern (both averaged over rotations and aligned with
    the plane XY).}
  \label{fig:comparison-patterns}
\end{figure*}

@@

Another interesting question is which sampling pattern is more powerful?
Our results shown in \cref{fig:comparison-patterns} clearly illustrate that
for isotropic structure all sampling techniques provide quite similar results.
The differences, of course, will be significant for isotropic cases and
directional 3-point CFs calculations. Again, the clear winner here is not obvious.
Right-triangular pattern allows to systematically sample the structure with
predefined direction. On the other hand, equilateral triangle allows to
minimize the volume of computed functions and represent them as single curve,
similar to 2-point functions. Random pattern could be useful for rigorous bounds.
So, the better approach can depend on the problem at hand. Nonetheless,
the direct comparison of patterns can be performed only on the basis of
information content -- something we discuss in the next paragraph.
